Hierarchically Supervised Latent Dirichlet Allocation

### 2011-06-01 paper/experiments.tex
Modified the experimental section (data and preprocessing) Tightened intro Added an accent on my name

%\begin{itemize}
%\item Describe Data for both Experiments
%\item Describe the Four different conditions
%\item Describe how they are all evaluated, with and without hierarchy
%\end{itemize}
In the section we describe the application of HSLDA for prediction
in two hierarchically structured domains. Firstly, we describe using
discharge summaries to predict diagnoses, encoded as ICD-9 codes.
Discharge summaries are documents that are authored by clinicians
to summarize the course of a hospitalization. ICD-9 codes are used
mainly for billing purposes to indicate the conditions for which a
patient was treated. Secondly, we describe using Amazon.com product
descriptions to predict product categories.

@@

Our data set was gathered from the clinical data warehouse of NewYork-Presbyterian
Hospital. The data consisted of free-text discharge summaries and
their respective ICD-9 codes. A discharge summary is a clinical report
prepared by a physician or other health professional at the conclusion
of a hospital stay or series of treatments. The note outlines the
patient's chief complaint, diagnostic findings, therapy administered,
patient's response to the chosen therapy, the treatment plan and the
recommendations upon discharge. The ICD-9 codes used to structure
the discharge summary data are part of a controlled terminology which
is the international standard diagnostic classification for epidemiological,
health management, and clinical purposes. The ICD-9 codes are organized
in a rooted-tree structure, with each edge representing an is-a relationship
between parent and child, such that the parent diagnosis subsumes
the child diagnosis. For example, the code representing {}``Pneumonia
due to adenovirus'' is a child of the code representing {}``Viral
pneumonia'' where the former is a type of the latter. In the hospital,
ICD-9 codes are generated manually by trained medical coders, who
review all the information in the discharge summary.

The text of the discharge summaries were pre-processed such that each
document would be represented as counts over a 10,000 word vocabulary.
The Natural Language Toolkit was used to tokenize the text. A vocabulary
was identified by first sorting terms based on a global term frequency-inverse
document frequency measure. The top 10,000 words which were not identifying
in some way (a name, place, or identifying number) were selected for
inclusion in the vocabulary.

For each hospitalization there are usually several ICD-9 codes assigned
for billing purposes. These codes are known to be quite specific but
not very sensitive \citep{Birmetal2005}. Regardless of that fact,
this is one of the only sources for information on patient diagnoses
aside from the free text. %Aside from prediction, one of the goals is to compare the sensitivity of predictions from the HSLDA model in comparison to the codes in a case where a test closer to ground truth is available. For this we will compare whether predictions for the ICD-9 code associated with anemia are better predicted by HSLDA or by the ICD-9 codes. Anemia was chosen because hemoglobin values are readily available and the definition of anemia according the World Health Organization is approximately 12.5, with a threshold of 12 for women and 13 for men [citation].

We worked within the guidelines of the Health Insurance Portability
and Accountability Act (HIPAA), which protects patient privacy and
the security of potentially identifying medical material, known as
personal health information (PHI). HIPAA covers any information within
a medical record that was created, used, or disclosed during the course
of providing a health care service and that can be used to identify
an individual. This study was approved by the Institutional Review
Board.

@@

Data for these experiments were obtained partially from the Stanford
Network Analysis Platform (SNAP) Amazon product metadata dataset \citep{SNAP}
and partially directly from the the Amazon.com website \citep{AMAZON}.
The product ID's and categorizations were obtained from the SNAP dataset
and the product descriptions were obtained directly from the website.

We were able to deduce the structure of the hierarchy for the Amazon.com
products directly since all ancestors in the hierarchy were included
with each category label. For example, {}``DVD / Genres / Science
Fiction \& Fantasy / Classic Sci-Fi'' is a single product category
for the DVD, {}``The Time Machine''. Each product was labeled with
multiple categories.

The vocabulary for this experiment was created by including the most
frequent 30K words omitting stopwords.

>>>>

%\begin{itemize}
%\item Describe Data for both Experiments
%\item Describe the Four different conditions
%\item Describe how they are all evaluated, with and without hierarchy
%\end{itemize}

We experimented with HSLDA for prediction in two domains: predicting medical
diagnosis codes from hospital discharge summaries and predicting product
categories from Amazon.com product descriptions.

%In the section we describe the application of HSLDA for prediction
%in two hierarchically structured domains. Firstly, we describe using
%discharge summaries to predict diagnoses, encoded as ICD-9 codes.
%Discharge summaries are documents that are authored by clinicians
%to summarize the course of a hospitalization. ICD-9 codes are used
%mainly for billing purposes to indicate the conditions for which a
%patient was treated. Secondly, we describe using Amazon.com product
%descriptions to predict product categories.

@@

Discharge summaries are authored by clinicians to summarize the course of a
hospitalized patient. The summaries typically contain a record of the patient
complains, findings and diagnoses, along with treatment and hospital course.
For each admission trained medical coders review the information in the
discharge summary and assign a series of diagnoses codes. Coding follows the
ICD-9-CM controlled terminology, an international diagnostic classification for
epidemiological, health management, and clinical
purposes.\footnote{http://www.cdc.gov/nchs/icd/icd9cm.htm} As such, the ICD-9
codes constitute a labeling of a patient's diagnoses based on a discharge
summary. The ICD-9 codes are organized in a rooted-tree structure, with each
edge representing an is-a relationship between parent and child, such that the
parent diagnosis subsumes the child diagnosis. For example, the code for
{}``Pneumonia due to adenovirus'' is a child of the code for {}``Viral
pneumonia,'' where the former is a type of the latter.  It is worth noting that
the coding can be noisy. Human coders sometimes disagree~\cite{Challenge07},
tend to be more specific than sensitive in their
assignments~\cite{Birmetal2005}, and sometimes make
mistakes~\cite{FarzandipourEtAl10}.

The task of automatic ICD-9 coding has been investigated in the clinical
domain. Much of the work was triggered by the 2007 medical NLP community
challenge~\cite{Challenge07}. The data in the challenge, however, differs from 
ours in its scope. The datasets were smaller (1,000 training and 1,000 testing
documents) and focused on radiology reports with a restricted number of ICD-9
codes (45 of them, compared to 7K+ in our dataset). Methods ranged from manual
rules to online learning~\cite{Crammer2007,Goldstein2007,Farkas2008}.
Other work had leveraged larger datasets and experimented with K-nearest
neighbor, Naive Bayes, support vector machines, Bayesian Ridge Regression, as
well as simple keyword mappings, all with promising
results~\cite{LarkeyCroft95,RibeiroNeto2001,PakhomovEtAl06,Ruch2008,Lita2008}.

Our dataset was gathered from the clinical data warehouse of
NewYork-Presbyterian Hospital. It consists of 6,000 discharge summaries and
their associated ICD-9 codes (7,298 distinct codes overall), representing all
the discharges from the hospital in 2009. Summaries have 8.39 associated ICD-9
codes on average (std dev=5.01) and contain an average of 536.57 terms after
preprocessing (std dev=300.29). We split our dataset into 5,000 discharge
summaries for training and 1,000 for testing.

The text of the discharge summaries was tokenized with
NLTK.\footnote{http://www.nltk.org} Vocabulary was determined as the top 10,000
tokens with highest document frequency (exclusive of names,
places and other identifying numbers). Each discharge summary is thus
represented as counts over the 10,000-word vocabulary. The study was approved
by the Institutional Review Board and follows HIPAA (Health
Insurance Portability and Accountability Act) privacy guidelines.

%For each hospitalization there are usually several ICD-9 codes assigned
%for billing purposes. These codes are known to be quite specific but
%not very sensitive \citep{}. Regardless of that fact,
%this is one of the only sources for information on patient diagnoses
%aside from the free text. %Aside from prediction, one of the goals is to compare the sensitivity of predictions from the HSLDA model in comparison to the codes in a case where a test closer to ground truth is available. For this we will compare whether predictions for the ICD-9 code associated with anemia are better predicted by HSLDA or by the ICD-9 codes. Anemia was chosen because hemoglobin values are readily available and the definition of anemia according the World Health Organization is approximately 12.5, with a threshold of 12 for women and 13 for men [citation].

@@

In this experiment, we look at product descriptions and their
categorizations according to a product hierarchy. Product ID's
and categorizations were obtained from the Stanford Network Analysis Platform
(SNAP) dataset~\citep{SNAP}. Product descriptions were crawled from the
amazon.com website directly. We were able to deduce the structure of the
hierarchy for the Amazon.com products directly since all ancestors in the
hierarchy were included with each category label. For example, {}``DVD / Genres
/ Science Fiction \& Fantasy / Classic Sci-Fi'' is a single product category 
for the DVD, {}``The Time Machine.''

Our dataset contains 15,130 product descriptions for training and 1,000
for testing. The product descriptions are shorter than the discharge summaries
(91.89 terms on average, std dev=53.08). Overall, there are 2,691 unique codes.
Products are assigned on average 9.01 codes (std dev=4.91). The vocabulary
consists of the most frequent 30,000 words omitting stopwords.

### 2011-06-01 paper/introduction.tex
Modified the experimental section (data and preprocessing) Tightened intro Added an accent on my name

There exist many sources of unstructured data that have been partially or completely categorized by human editors.  In this paper we focus on unstructured textual data that has been, at least in part, manually categorized.  Examples include but are not limited to webpages and curated hierarchical directories of the same \citep{DMOZ}, product descriptions and catalogs (e.g.~\citep{AMAZON} as available from \citep{SNAP}), and patient hospital treatment transcripts and  codes applied to them for bookkeeping and insurance purposes (e.g.~hospital discharge records with International Classification of Disease 9th Revision,
Clinical Modification (ICD-9-CM) codes assigned \cite{ICD9}).  In this work we show how to combine these two sources of information using a single model that allows one to automatically categorize new text documents, suggest labels that might be inaccurate, compute improved similarities between documents for information retrieval purposes, and more.  The models and techniques that we develop in this paper are applicable in other domains as well, namely, any unstructured representations of data that have been hierarchically classified (e.g.~image catalogs with bag-of-feature image representations).

>>>>

There exist many sources of unstructured data that have been partially or
completely categorized by human editors.  In this paper we focus on
unstructured textual data that has been, at least in part, manually
categorized.  Examples include but are not limited to webpages and curated
hierarchical directories of the same \citep{DMOZ}, product descriptions and
catalogs~\citep{SNAP}, and patient records and their associated billing codes.
In this work we show how to combine these two sources of information using a
single model that allows one to automatically categorize new text documents,
suggest labels that might be inaccurate, compute improved similarities between
documents for information retrieval purposes, and more. The models and
techniques that we develop in this paper are applicable in other domains as
well, namely, any unstructured representations of data that have been
hierarchically classified (e.g.~image catalogs with bag-of-feature image
representations).
